\Large \bf Relationship of Weather and Maize Yields in Kenya

### 2018-11-14 writing/draft1/manuscript11.tex
adding

\begin{itemize}
\item The yields seem to be more responsive to weather on west than on east
\end{itemize}
\normalsize
\FloatBarrier
\pagebreak
This aim of this paper is to contribute to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security in Kenya.

@@

>>>>

\begin{itemize}
\item The yields seem to be more responsive to weather on west than on east
\end{itemize}
\normalsize
\FloatBarrier
\pagebreak
The aim of this paper is to contribute to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security in Kenya.

@@

\textcolor{red}{Then maybe something about agriculture in Kenya. for example:}
 \\
 \cite{Kabubo2015} \\
 \cite{WorldBank2015} \\
  \cite{Lesk2016}: Global, but estimate of national cereal production losses from extreme weather disasters
 \\

\textcolor{red}{Then maybe something about proactive approach:}

### 2018-11-14 writing/draft1/manuscript12.tex
adding

\begin{abstract}
\noindent We explore an unprecedented dataset of almost $6,000$ observations to identify main predictors of climate knowledge, climate risk perception and willingness to pay for climate change mitigation. Among nearly $70$ potential explanatory variables we detect the most important ones using multisplit lasso estimator. Importantly, we test significance of individuals' preferences about time, risk and equity. Our study is innovative as these behavioural characteristics were recorded by including experimental methods into a live sample survey. This unique way of data collection combines advantages of survey and experiments. The most important predictors of environmental attitudes are numeracy, cognitive ability, ideological world-view and inequity aversion. \\
\end{abstract}

\noindent \textbf{JEL classification:} Q54, Q58, D80\\
\noindent \textbf{Keywords:} Climate change, climate knowledge, climate policy, lasso, risk perception, willingness to pay\\

@@

\begin{itemize}
\item The yields seem to be more responsive to weather on west than on east
\end{itemize}
\normalsize
\FloatBarrier
\pagebreak
The aim of this paper is to contribute to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security in Kenya.

@@

Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}. During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia. Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}. \cite{Trenberth2014} argue that as a consequence of global warming, dry areas have strong tendency to get drier while wet areas are getting wetter.

@@

\textcolor{red}{Then maybe something about proactive approach:}

Recent literature has promoted shifting from reactive to proactive approach in disaster risk reduction. Forecast based financing has been promoted as it can avoid significant disaster losses \citep{Nicholson2017}. It has been shown that the investment into disaster risk reduction had usually been outweighed by avoided losses \citep{Mechler2005}.

@@

\textcolor{blue}{\textbf{Add more studies. Then separate somehow into groups of different topics..}}

@@

Prior we started  our research, we had to answer the following questions: How do we measure food security? And how do we measure drought? To answer the first questions, the measure which has been used as a proxy for food security in the literature are as follows:

\vspace{2cm}

>>>>

\begin{abstract}
\noindent This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security. 
\\
\end{abstract}

\noindent \textbf{Keywords:} Climate change, climate knowledge, climate policy, lasso, risk perception, willingness to pay\\

@@

\begin{itemize}
\item The yields seem to be more responsive to weather on west than on east
\end{itemize}
\normalsize
\FloatBarrier
\pagebreak

It is well known that drought and extreme heat waves affect agriculture and agricultural production mostly in negative ways (\citealp{Deschenes2007Ric,Lesk2016,Mehrabi2017}) \textcolor{blue}{I can go into details about these studies if not long enough}. As a consequence of climate change, the situation is likely to get worse as dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014}. Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}. During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia. Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.

Drought has been one of the most serious problem in Kenya as well as in other Sub-Saharan countries.

@@

@@

Shifting from reactive to proactive approach in disaster risk reduction has been promoted in recent literature (\citealp{Mechler2005,IPCC2012ch1,Nicholson2017}). Forecast based financing has been recommended as it can avoid significant disaster losses \citep{Nicholson2017}. It has been shown that investment into disaster risk reduction had usually been outweighed by avoided losses \citep{Mechler2005}. Risk assessment is essential for disaster risk reduction \citep{IPCC2012ch1}. Therefore, the main purpose of this paper is to develop a model for assessing the risk of drought based on weather data.

@@

\textcolor{blue}{\textbf{Add more studies. Then separate somehow into groups of different topics..}}
\textcolor{green}{There has been a lot of literature about impacts of climate change on agriculture..Mendelsohn, Dinar, Dalfelt 2000 and 2009 and Seo, Mendelsohn, Dinar, Hassan, Kurukulasuriya 2008
}

@@

Prior we started  our research, we had to answer the following questions: How do we measure food security? And how do we measure drought? To answer the first questions, the measure which has been used as a proxy for food security in the literature are as follows:\\

\textcolor{red}{We were also considering to use food prices or export and import prices/retail prices, however, there were some problems...}
\vspace{2cm}

### 2018-11-14 writing/draft1/manuscript13.tex
adding

\section{Introduction}\label{Introduction}

@@

It is well known that drought and extreme heat waves affect agriculture and agricultural production mostly in negative ways (\citealp{Deschenes2007Ric,Lesk2016,Mehrabi2017}) \textcolor{blue}{I can go into details about these studies if not long enough}. As a consequence of climate change, the situation is likely to get worse as dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014}. Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}. During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia. Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.

Drought has been one of the most serious problem in Kenya as well as in other Sub-Saharan countries.

\textcolor{red}{something about droughts in Kenya, how they were recently and what were the effects, find refrences..see my folder literatureFS/references SSRP or the book East African Agriculture and Climate in the folder literatureFS} for example \citep{Nicholson2017}

\textcolor{red}{Then maybe something about agriculture in Kenya. for example:}
 \\
 \cite{Kabubo2015} \\
 \cite{WorldBank2015} \\
  \cite{Lesk2016}: Global, but estimate of national cereal production losses from extreme weather disasters
 \\

@@

An important group of studies analysing the relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).

@@

\textcolor{blue}{\textbf{Add more studies. Then separate somehow into groups of different topics..}}
\textcolor{green}{There has been a lot of literature about impacts of climate change on agriculture..Mendelsohn, Dinar, Dalfelt 2000 and 2009 and Seo, Mendelsohn, Dinar, Hassan, Kurukulasuriya 2008
}

@@

\subsection{Measures of Yield and climate}

@@

\textcolor{red}{We were also considering to use food prices or export and import prices/retail prices, however, there were some problems...}
\vspace{2cm}

@@

In the recent period, remote sensing data have been collected and used increasingly to monitor levels of greenness and closely related vegetation conditions. Based on these data, the vegetation condition index (VCI) has been developed for quantifying drought strength and severity \citep{KlischAtz2016}.

@@

\begin{itemize}

@@

\subsection{Linear Mixed Models}
\sloppy
Kenya consists of $47$ counties with semi-autonomous county governments  \citep{Barasa2017}. As a result of the high degree of county-level autonomy, the policies and regulations are likely to differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties. Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (or linear mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}. These types of models include both fixed affects and random effects. Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}. On the other hand, random effect are unobserved random variables. There are at least three benefits of treating a set of parameters as a random sample from some distribution. \textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9,KERYch12}).

@@

\subsection{Climate change risk perception}\label{ResPerc}

In this section we discuss our estimates of the models which explain individuals' perception of climate change risk. We focus on two measures of climate risk perception, in particular climate change seriousness perception and climate versus policy perception. We present the results of lasso and jackknife OLS with the climate seriousness perception as dependent variable in Table~\ref{Climcare435}. Three predictors were selected, in particular gender, climate knowledge, and degree of agreement with redistribution of income by government. In this case, the effect of being male is negative. This is mostly consistent with results of previous research which typically finds women to take climate risk more seriously than men (\citealp{WHITMARSH2011, McCright2010, Kahan2007}). As we can see in Table~\ref{Climcare435}, degree of agreement with income redistribution affects climate change seriousness perception positively as the base category is 'Strongly disagree'. This is in agreement with previous literature as we consider the degree of agreement with income redistribution as an indicator of political and ideological world-view, which was found to be significantly correlated with climate concern by large number of previous studies (e.g. \citealp{Leiserowitz2013, Kahan2012, WHITMARSH2011}).

We will comment on the significant effects of climate knowledge at the end of Section~\ref{ResPerc}.

{\centering
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{Climate change seriousness perception:} Multisplit lasso and jackknife OLS}}
% soubor LassoClimcare43 Climcare435.R

\label{Climcare435} 
\centering
\begin{small}
\begin{tabular}{lclrcl} 
\hline \vspace{-0.2cm} \\
  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\

\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
 \hline 
\hline
\\
\vspace{-0.2cm}Gender = male&$0.0002$&$^{***}$&$-0.3658$&$4.45\times 10^{-6}$&$^{***}$\\
  \\
\vspace{-0.2cm}Climate knowledge&$1.0000$& &$0.1380$&$1.0000$&\\
  \\
\vspace{-0.2cm}Climate knowledge - squared&$<2.00\times 10^{-8}$&$^{***}$&$-0.0548$&$0.0209$&$^{*}$\\
  \\
 Redistribution of income:&\multirow{2}{*}{$1.0000$}& &\multirow{2}{*}{$0.1819$}&\multirow{2}{*}{$1.0000$}&\\%-1
      \hspace{0.6cm}disagree\tnote{a}&& &&&\\%-1
    \\
  \vspace{-0.2cm}Redistribution of income: neutral\tnote{a}&$1.0000$& &$0.2789$&$0.8251$&\\%0
    \\
  \vspace{-0.2cm}Redistribution of income: agree\tnote{a}&$<2.00\times 10^{-8}$&$^{***}$&$0.8343$&$8.58\times 10^{-8}$&$^{***}$\\%1
    \\
  Redistribution of income:& \multirow{2}{*}{$<2.00\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$1.0828$}&\multirow{2}{*}{$<2.00\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
\hspace{0.6cm}strongly agree\tnote{a}&& &&&\\%2
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
   \item[~]\textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
  \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.

 \item[a] Degree of agreement with the following statement: 'Government should redistribute income from the better off to those who are less well off.' The base category is 'Strongly disagree'.
     \end{adjustwidth}
 \singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

\hspace{1cm}

>>>>

\section{Introduction  and literature research}\label{Introduction}

@@

Effects of drought and extreme heat waves on agriculture, which are mostly negative, have been well documented (\citealp{Deschenes2007Ric,Lesk2016,Mehrabi2017, schwalm2017}). \textcolor{blue}{I can go into details about these studies if not long enough}. Agricultural drought and soil moister are important factors for plant health but also for land-atmospheric feedback and temperature variability \citep{nicolai2017}. As a consequence of climate change, the situation about droughts is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}. Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}. During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia. Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.

Frequent occurrence of severe drought has been a major problem in Kenya as well as in other Sub-Saharan countries \citep{Nicholson2017}. Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment. In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}. Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.

Large literature has focused on relationship of agricultural production and climate.

\textcolor{red}{Then maybe something about agriculture in Kenya. for example:}
 \\
 \cite{Kabubo2015} \textcolor{green}{have a look at this study. Looks like very much similar to what I want to estimate. Kenya, panel, fixed and random effects...}\\
 \cite{WorldBank2015} \\
  \cite{Lesk2016}: Global, but estimate of national cereal production losses from extreme weather disasters
 \\

Previous studies focused on responses to drought related disasters or emergency situations include \cite{Sandstorm2017}, who question whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia. They analyse causes of the food crisis identified in the humanitarian appeal documents. Majority of these documents consider food availability and food production as a major cause of food crisis. \cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.  \cite{Sandstorm2017} also find that a large share of humanitarian response budget is focused on emergency food aid in comparison to the share focused on interventions to build resilience. They suggest to increase the budget share focused on agricultural and livestock production to build resilience.

@@

An important group of studies that analyse relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).

@@

\textcolor{blue}{\textbf{Add more studies. Then separate somehow into groups of different topics..}}
\textcolor{green}{It is worth mentioning at this place that large literature has investigated impacts of climate change on agriculture..worldwide or in Kenya??Mendelsohn, Dinar, Dalfelt 2000 and 2009 and Seo, Mendelsohn, Dinar, Hassan, Kurukulasuriya 2008
}

@@

\subsection{Measures of yield and climate}

@@

\textcolor{red}{We were also considering to use food prices or export and import prices/retail prices, however, there were some problems...maybe some references to those who also considered food prices??}
\vspace{2cm}

@@

In the recent period, remote sensing data have been collected and used increasingly to monitor levels of greenness and closely related vegetation conditions \citep{nicolai2017}. Based on these data, various measures such as the vegetation condition index (VCI) have been developed and used for quantifying drought strength and severity \citep{KlischAtz2016}.

@@

\textcolor{blue}{maybe also describe the two rainy seasons and why we created the monthly variables? or maybe this in intro?}
\begin{itemize}

@@

\subsection{Linear mixed models}
\sloppy
Kenya consists of $47$ counties with semi-autonomous county governments  \citep{Barasa2017}. As a result of the high degree of county-level autonomy, the policies and regulations are likely to differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties. Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (or linear mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}. These types of models include both fixed affects and random effects. Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}. On the other hand, random effect are unobserved random variables. There are at least three benefits of treating a set of parameters as a random sample from some distribution. \textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9,KERYch12}).

@@

### 2018-11-14 writing/draft1/manuscript14.tex
adding

Effects of drought and extreme heat waves on agriculture, which are mostly negative, have been well documented (\citealp{Deschenes2007Ric,Lesk2016,Mehrabi2017, schwalm2017}). \textcolor{blue}{I can go into details about these studies if not long enough}. Agricultural drought and soil moister are important factors for plant health but also for land-atmospheric feedback and temperature variability \citep{nicolai2017}. As a consequence of climate change, the situation about droughts is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}. Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}. During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia. Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.

Frequent occurrence of severe drought has been a major problem in Kenya as well as in other Sub-Saharan countries \citep{Nicholson2017}. Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment. In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}. Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.

Large literature has focused on relationship of agricultural production and climate.

@@

Previous studies focused on responses to drought related disasters or emergency situations include \cite{Sandstorm2017}, who question whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia. They analyse causes of the food crisis identified in the humanitarian appeal documents. Majority of these documents consider food availability and food production as a major cause of food crisis. \cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.  \cite{Sandstorm2017} also find that a large share of humanitarian response budget is focused on emergency food aid in comparison to the share focused on interventions to build resilience. They suggest to increase the budget share focused on agricultural and livestock production to build resilience.

>>>>

Consequences of drought and extreme heat waves on agriculture have been well documented (\citealp{Deschenes2007Ric,RicardianBello,Lesk2016,Mehrabi2017, schwalm2017}). \textcolor{blue}{I can go into details about these studies if not long enough}. Agricultural drought and soil moister are important factors for plant health but also for land-atmospheric feedback and temperature variability \citep{nicolai2017}.

Frequent occurrence of severe drought has been a major problem in Kenya as well as in other Sub-Saharan countries \citep{Compendium,WorldBank2015, Nicholson2017}. Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment. In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}. Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.  As a consequence of climate change, the situation is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}. Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}. During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia. Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.

Large literature has focused on relationship of agricultural production and climate in Sub-Saharan Africa. A frequent finding is that increase in temperature affects crop revenue negatively while increase in precipitation has positive effects.  \cite{RicardianBello} and  \cite{Ochieng2016} found these results by the means of Ricardian analysis. \cite{Ochieng2016} focused on Kenya while \cite{RicardianBello} address the situation on Niger. \cite{WorldBank2015} \textcolor{red}{identify and assess risks to agriculture in Kenya.}

Other studies investigate the relationship of climate and yields in the context of climate change. \cite{kabubo2007} argue that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial. Furthermore, fall and summer precipitation is positively correlated with net revenue. \cite{KMendelsohn2008} and \cite{SeoMendelsohn} investigate the distribution of climate change impacts on agriculture in Africa. They argue that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful.
\cite{Mendelsohn2008} reviewed several papers on effects of climate change on agriculture in developing countries and confirms that tropical and sub-tropical counties are much more sensitive to climate change than temperate agriculture. Another meta-study of climate change impacts of crop productivity in Africa in South Asia was conducted by \cite{Knox2012} who conclude that the projected mean change in yield is -$8\%$ by the 2050s in both regions.

@@

Previous studies focused on responses to drought related disasters or emergency situations include \cite{Sandstorm2017}, who questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia. They analysed causes of the food crisis identified in the humanitarian appeal documents. Majority of these documents consider food availability and food production as a major cause of food crisis. \cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.  \cite{Sandstorm2017} also find that a large share of humanitarian response budget is focused on emergency food aid in comparison to the share focused on interventions to build resilience. They suggest to increase the budget share focused on agricultural and livestock production to build resilience.
